Drug sensitivity prediction with normal inverse Gaussian shrinkage informed by external data

### 2020-04-15 docs/supplement1.Rnw
commit before push for bSEM analysis 4
changed name of project

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
knit_hooks$set(document=function(x) {
  sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed=TRUE)})
@

@@

	\section{Hyperparameter settings}\label{sec:hyperparameters}
	For the NIG shrinkage weights prior in Figure \ref{md-fig:dens_kappa}a the 
	following hyperparameter settings (in terms of the inverse Gaussian prior) 
	were used: $\lambda=1$, $\phi=10$ (solid),
	$\lambda=0.1$, $\phi=10$ (dashed), and $\lambda=1$, $\phi=1$ (dotted).
	The Student`s $t$ prior on the regression parameters $\beta_{jd}$ arises if
	$\gamma_{jd}^2 \sim \Gamma^{-1}(a,b)$, with $\Gamma^{-1}(a,b)$ an inverse
	Gamma distribution with shape $a$ and scale $b$. Figure 
	\ref{md-fig:dens_kappa}b is then obtained by setting: $a=0.9$, $b=1.3$ 
	(solid), $a=1$, $b=4$ (dashed), and $a=1$, $b=0.2$ (dotted).
	The lasso prior results from an exponential prior on $\gamma_{jd}^2$. To
	create Figure \ref{md-fig:dens_kappa}c the following settings were
	used for rate $\lambda$: $\lambda=1$ (solid), $\lambda=0.2$ (dashed), and 
	$\lambda=0.1$ (dotted).

	\section{Jeffrey's prior}\label{sec:jeffreysprior}
	We parametrise the data with a Gaussian distribution with unknown mean and 
	variance, i.e., $y_i \sim \mathcal{N} (\mu, \sigma^2)$. This gives us the
	following Hessian (matrix of second derivatives) of the log density:
	$$
	\mathbf{H}(\log f(y_i | \mu, \sigma^2)) = \begin{bmatrix}
	-\frac{1}{\sigma^2} & y_i - \mu \\
	y_i - \mu & \frac{1}{2\sigma^4} - \frac{(y_i - \mu)^2}{\sigma^6}.
	\end{bmatrix}
	$$
	Jeffrey's prior is now found as:
	$$
	\pi(\mu, \sigma^2) \propto |\mathcal{I}(\mu, \sigma^2)|^{1/2} = 
	|-\E (\mathbf{H})|^{1/2} = \begin{vmatrix}
	\frac{1}{\sigma^2} & 0 \\
	0 & \frac{1}{2\sigma^4}
	\end{vmatrix}^{1/2} \propto 1/\sigma^3,
	$$
	where $\mathcal{I}(\theta)$ denotes the Fisher information of $\theta$.

@@

  \section{Evidence lower bound}\label{sec:elbo}
  The evidence lower bound that is maximised during the iterations is: 
  $\text{ELBO}^{(l)}=\text{ELBO}_d^{(l)}$, with:
  \begin{align*}
    \text{ELBO}&^{(l)}_d \\
    & = -\frac{n+p+1}{2} \log \pi + \frac{p+1-n}{2}\log 2 + \frac{n+1}{2} + p
    + \log \Gamma \(\frac{n + p + 1}{2} \) \\
    & - \frac{n + p + 1}{2} \log \zeta_d^{(l)} + 
    \frac{1}{2} \log |\bSigma_d^{(l)}| - 
    \frac{n+p+1}{4\zeta_d^{(l)}}  \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(l)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(l)}) \\ 
    & + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} +
    g_{d}^{(l)} \trace \[ \diag (b_{jd}^{(l)}) \bm{\Sigma}_d^{(l)}\] + 
    g_{d}^{(l)} (\bm{\mu}_d^{(l)}) \tr \diag (b_{jd}^{(l)}) 
    \bm{\mu}_d^{(l)}\bigg] + p \log \lambda_{\text{feat}}^{(l)} \\
    & + \frac{p + 3}{4} \log \lambda_{\text{drug}}^{(l)} + 
    \frac{\eta_d^{(l)} - \lambda_{\text{drug}}^{(l)}}{2} g_d^{(l)} + 
    \sum_{j=1}^p \frac{\delta_{jd}^{(l)} - \lambda_{\text{feat}}^{(l)}}{2} 
    b_{jd}^{(l)} + \frac{\lambda_{\text{drug}}^{(l)}}{\chi_d^{(l)}} +
    \sum_{j=1}^p \frac{\lambda_{\text{feat}}^{(l)}}{\phi_{jd}^{(l)}} + \\
    & - \frac{p + 1}{2} \log \chi_d^{(l)} - \sum_{j=1}^p \log \phi_{jd}^{(l)}
    -\frac{p+1}{4} \log \eta_d^{(l)} - 
    \frac{1}{2} \sum_{j=1}^p \log \delta_{jd}^{(l)} \\
    & + \log K_{\frac{p+1}{2}}\(\sqrt{\eta_{d}^{(l)}
    \lambda_{\text{drug}}^{(l)}}/
    \chi_{d}^{(l)}\) + \sum_{j=1}^p \log K_{1}\(\sqrt{\delta_{jd}^{(l)}
    \lambda_{\text{feat}}^{(l)}}/\phi_{jd}^{(l)}\).
  \end{align*}

  \section{Efficient computation}\label{sec:computation}
  The empirical Bayes updates require the quantities
  $e_{jd}$, $b_{jd}$, $g_d$, and $f_d$, which in turn depend on the quantities 
  $\diag(\bSigma_d)$, $\bmu_{d}$, $\y_d \tr \X \bmu_d$, 
  $\trace(\X \tr \X \bSigma_d)$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$.
  Let $\mathbf{H}_d=g_d \diag[(b_{jd}^{(h)})^{-1}]$. Then, the first two
  quantities are efficiently calculated as:
  \begin{subequations}\label{eq:complexity1}
    \begin{align}
      \diag (\bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} 
      \H_d^{(h)} - (a_d^{(h)})^{-1} \{ [ \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} ] \circ
      (\H_d^{(h)} \X \tr ) \} \mathbf{1}_{n \times 1}, \\
      \bmu_d^{(h+1)}) & = \H_d^{(h)} \X \tr \y_d - \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr \y_d.
    \end{align}
  \end{subequations}
  The quantities $\y_d \tr \X \bmu_d$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$
  are easily calculated from the first two. For the remaining one we have:
  \begin{align}\label{eq:complexity2}
    \trace(\X \tr \X \bSigma_d^{(h+1)})  = 
    & (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \H_d^{(h)} \X \tr 
    \mathbf{1}_{n \times 1} \nonumber \\  
    & - (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \{ [  \H_d^{(h)} \X \tr 
    ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr ] \circ 
    \X \tr \} \mathbf{1}_{n \times 1}.
  \end{align}
  Both (\ref{eq:complexity1}) and (\ref{eq:complexity2}) are operations of
  complexity $\mathcal{O}(pn^2)$.

@@

  We let $n_1=p/2$ and $n_2=(p-1)/2$ and use the well-known recursive relation:
  $$
  K_{\alpha}(x) = K_{\alpha - 2}(x) + \frac{2(\alpha - 1)}{2} K_{\alpha- 1}(x),
  $$
  to rewrite the ratio:
  \begin{align*}
    \frac{K_{\frac{p - 1}{2}}(x)}{K_{\frac{p + 1}{2}}(x)} = 
    \begin{cases}
      \Big( \dots \Big( \( 1 + \frac{2 \cdot 1 - 1}{x} \)^{-1} + 
      \frac{2 \cdot 2-1}{x}\Big)^{-1} + \dots + 
      \frac{2 \cdot n_1 - 1}{x}\Big)^{-1}, & \text{for } p \text{ even}, \\
      \Big( \dots \Big( \( \frac{K_0(x)}{K_1(x)} + \frac{2}{x} \cdot 1 \)^{-1} +
      \frac{2}{x} \cdot 2 \Big)^{-1} + \dots + \frac{2}{x} \cdot n_2 \Big)^{-1},
      & \text{for } p \text{ odd}.
    \end{cases}
  \end{align*} 
  The ratio $K_0(x)/K_1(x)$ is well-behaved, so that it may be computed
  without numerical issues.

@@

  Next, we note that through Jensen's inequality 
  $e_{jd}^{-1}=\E(\gamma_{jd}^{2})^{-1} \leq \E(\gamma_{jd}^{-2})=b_{jd}$,
  we may lowerbound the summands in left-hand side of (\ref{eq:inequality}):
  $$
  b_{jd} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} \geq
  e_{jd}^{-1} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1}.
  $$
  which we use to show that:
  \begin{align*}
    e_{jd}^{-1} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} & \geq 0 \\
    1 + \phi_{jd}^{-2} e_{jd}^2 - 2 \phi_{jd}^{-1} e_{jd} & \geq 0 \\
    \( 1 - \phi_{jd}^{-1} e_{jd} \)^2 & \geq 0.
  \end{align*}
  Similar reasoning holds for the drug scale parameter $\lambda_{\text{drug}}$.

  \section{Estimation of alternative prior}\label{sec:altprior}
  The hyperparameters $\bm{\alpha} = \begin{bmatrix} \bm{\alpha}_{\text{feat}}
  \tr & \bm{\alpha}_{\text{drug}} \tr \end{bmatrix} \tr$ in model 
  (\ref{md-eq:altprior}) are estimated by empirical Bayes maximisation of
  the marginal log likelihood:
  \begin{subequations}\label{eq:altmll}
    \begin{align}
      \log p (\bm{\alpha} | \mathbf{Y}) & = \sum_{d=1}^D
      \log \[ \int_{\bm{\beta}_d} \int_{\sigma^2_d} \mathcal{L} 
      (\mathbf{y}_d | \bm{\beta}_d, \sigma^2_d ) \pi(\bm{\beta}_d | \sigma^2_d) 
      \pi(\sigma^2_d) d\bm{\beta}_d d\sigma_d^2 \] \\
      & \propto - \sum_{d=1}^D \bigg[ \log \left|\I + \mathbf{X} 
      \diag \[ \exp\( \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} \tr 
      \right| \\
      & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\, + (n + 1) \log \( \mathbf{y}_d \tr 
      \left\{ \I + \mathbf{X} 
      \diag \[ \exp\( \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} 
      \tr\right\}^{-1} \mathbf{y}_d \) \bigg],
    \end{align}
  \end{subequations}
  where $\tilde{\mathbf{C}}_d = \begin{bmatrix} \mathbf{C}_d & 
  \mathbf{1}_{p \times 1} \cdot \mathbf{z}_d \tr \end{bmatrix}$, and
  $\mathbf{C}_d = \begin{bmatrix} \mathbf{c}_{1d} & \cdots & 
  \mathbf{c}_{pd} \end{bmatrix} \tr$.
  The first term in the sum of (\ref{eq:altmll}) is concave in 
  $\bm{\alpha}$, while the second term is convex in $\bm{\alpha}$. This renders
  the marginal log likelihood a sum of non-concave (as well as non-convex) 
  functions. General purpose optimisers, such as \texttt{optim} in \texttt{R} 
  are generally not well suited to maximise non-concave functions. Maximisation
  is therefore done through a minorise-maximisation (MM) algorithm, 
  similar to \cite{zeng_incorporating_2020}.

  The MM algorithm seeks to minorise the marginal log likelihood,
  $\log p(\bm{\alpha} | \mathbf{Y})$, at current
  parameter estimate $\bm{\alpha}^{(t)}$ with function 
  $g(\bm{\alpha} | \bm{\alpha}^{(t)})$, such that
  $\forall \bm{\alpha}: g(\bm{\alpha} | \bm{\alpha}^{(t)}) \leq 
  \log p(\bm{\alpha} | \mathbf{Y})$ 
  and $g(\bm{\alpha}^{(t)} | \bm{\alpha}^{(t)}) = 
  \log p(\bm{\alpha} | \mathbf{Y})$. Then, the sequence
  $$
  \bm{\alpha}^{(t + 1)} = \underset{\bm{\alpha}}{\argmax} 
  g(\bm{\alpha} | \bm{\alpha}^{(t+1)})
  $$
  converges to a local maximum of $f(\bm{\alpha})$. Generally, the minorisation
  function is chosen such that it is concave and easy to maximise. In our case,
  a suitable minorisation is:
  \begin{subequations}\label{eq:minor}
    \begin{align}
      g(\bm{\alpha} | \bm{\alpha}^{(t+1)}) = -\sum_{d=1}^D \bigg[
      & \diag \( \X \tr \mathbf{A}_d^{(t)} \X \) \tr \exp (\tilde{\mathbf{C}}_d
      \bm{\alpha}) \\
      & + \frac{n + 1}{\mathbf{y}_d \tr \mathbf{A}_d^{(t)} \mathbf{y}_d} 
      \mathbf{y}_d \tr \left\{ \I + \mathbf{X} \diag \[ \exp\( 
      \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} \tr \right\}^{-1} 
      \mathbf{y}_d\bigg],
    \end{align}
  \end{subequations}
  where $\mathbf{A}_d^{(t)} = \left\{ \I + \mathbf{X} \diag \[ \exp\( 
  \tilde{\mathbf{C}}_d \bm{\alpha}^{(t)} \) \] \mathbf{X} \tr \right\}^{-1}$.
  Note that (\ref{eq:minor}) is concave. However, it is slow to optimise. 
  In \cite{zeng_incorporating_2020}, another minorization is proposed for
  the second term:
  $$
  \mathbf{y}_d \tr \left\{ \I + \mathbf{X} \diag \[ \exp\( 
  \tilde{\mathbf{C}}_d \bm{\alpha} \) \] \mathbf{X} \tr \right\}^{-1} 
  \mathbf{y}_d = \underset{\delta}{\max} 
  - ||\mathbf{y}_d - \mathbf{X} \delta ||_2^2 - 
  \delta \tr \diag \[ \exp\(
  -\tilde{\mathbf{C}}_d \bm{\alpha} \) \] \delta.
  $$
  The extra parameter $\delta$ is estimated by any standard ridge regression
  package, such as \texttt{glmnet} in \texttt{R}.

>>>>

% set options
<<settings, include=FALSE, echo=FALSE>>=
opts_knit$set(root.dir="..", base.dir="../figs")
opts_chunk$set(fig.align='center', fig.path="../figs/", 
               echo=FALSE, cache=FALSE, message=FALSE, fig.pos='!ht')
knit_hooks$set(document=function(x) {
  sub('\\usepackage[]{color}', '\\usepackage{xcolor}', x, fixed=TRUE)})
options(knitr.kable.NA="", digits=3)
@

@@

	\section{Hyperparameter settings}\label{sec:hyperparameters}
	For the NIG shrinkage weights prior in Figure \ref{md-fig:dens_kappa}a the 
	following hyperparameter settings (in terms of the inverse Gaussian prior) 
	were used: $\lambda=1$, $\phi=10$ (solid),
	$\lambda=0.1$, $\phi=10$ (dashed), and $\lambda=1$, $\phi=1$ (dotted).
	The Student`s $t$ prior on the regression parameters $\beta_{jd}$ arises if
	$\gamma_{jd}^2 \sim \Gamma^{-1}(a,b)$, with $\Gamma^{-1}(a,b)$ an inverse
	Gamma distribution with shape $a$ and scale $b$. Figure 
	\ref{md-fig:dens_kappa}b is then obtained by setting: $a=0.9$, $b=1.3$ 
	(solid), $a=1$, $b=4$ (dashed), and $a=1$, $b=0.2$ (dotted).
	The lasso prior results from an exponential prior on $\gamma_{jd}^2$. To
	create Figure \ref{md-fig:dens_kappa}c the following settings were
	used for exponential rate $\lambda$: $\lambda=1$ (solid), $\lambda=0.2$ 
	(dashed), and $\lambda=0.1$ (dotted).

	\section{Jeffrey's prior}\label{sec:jeffreysprior}
	We parametrise the data with a Gaussian distribution with unknown mean and 
	variance, i.e., $y_i \sim \mathcal{N} (\mu, \sigma^2)$. This results in the
	following Hessian (matrix of second derivatives) of the log density:
	$$
	\mathbf{H}(\log f(y_i | \mu, \sigma^2)) = \begin{bmatrix}
	-\frac{1}{\sigma^2} & y_i - \mu \\
	y_i - \mu & \frac{1}{2\sigma^4} - \frac{(y_i - \mu)^2}{\sigma^6}.
	\end{bmatrix}
	$$
	Jeffreys prior \cite[]{jeffreys_invariant_1946} is now found as:
	$$
	\pi(\mu, \sigma^2) \propto |\mathcal{I}(\mu, \sigma^2)|^{1/2} = 
	|-\E (\mathbf{H})|^{1/2} = \begin{vmatrix}
	\frac{1}{\sigma^2} & 0 \\
	0 & \frac{1}{2\sigma^4}
	\end{vmatrix}^{1/2} \propto 1/\sigma^3,
	$$
	where $\mathcal{I}(\theta)$ denotes the Fisher information of $\theta$.

@@

  \section{Evidence lower bound}\label{sec:elbo}
  The evidence lower bound that is maximised during the iterations is: 
  $\text{ELBO}^{(l)}=\sum_{d=1}^D \text{ELBO}_d^{(l)}$, with:
  \begin{align*}
    \text{ELBO}&^{(l)}_d \\
    & = -\frac{n+p+1}{2} \log \pi + \frac{p+1-n}{2}\log 2 + \frac{n+1}{2} + p
    + \log \Gamma \(\frac{n + p + 1}{2} \) \\
    & - \frac{n + p + 1}{2} \log \zeta_d^{(l)} + 
    \frac{1}{2} \log |\bSigma_d^{(l)}| - 
    \frac{n+p+1}{4\zeta_d^{(l)}}  \bigg[\mathbf{y}_d \tr \mathbf{y}_d -
    2 \mathbf{y}_d \tr \X \bm{\mu}_d^{(l)} + 
    \trace ( \X \tr \X \bm{\Sigma}_d^{(l)}) \\ 
    & + (\bm{\mu}_d^{(l)}) \tr \X \tr \X \bm{\mu}_d^{(l)} +
    g_{d}^{(l)} \trace \[ \diag (b_{jd}^{(l)}) \bm{\Sigma}_d^{(l)}\] + 
    g_{d}^{(l)} (\bm{\mu}_d^{(l)}) \tr \diag (b_{jd}^{(l)}) 
    \bm{\mu}_d^{(l)}\bigg] + p \log \lambda_{\text{feat}}^{(l)} \\
    & + \frac{p + 3}{4} \log \lambda_{\text{drug}}^{(l)} + 
    \frac{\eta_d^{(l)} - \lambda_{\text{drug}}^{(l)}}{2} g_d^{(l)} + 
    \sum_{j=1}^p \frac{\delta_{jd}^{(l)} - \lambda_{\text{feat}}^{(l)}}{2} 
    b_{jd}^{(l)} + \frac{\lambda_{\text{drug}}^{(l)}}{\chi_d^{(l)}} +
    \sum_{j=1}^p \frac{\lambda_{\text{feat}}^{(l)}}{\phi_{jd}^{(l)}} + \\
    & - \frac{p + 1}{2} \log \chi_d^{(l)} - \sum_{j=1}^p \log \phi_{jd}^{(l)}
    -\frac{p+1}{4} \log \eta_d^{(l)} - 
    \frac{1}{2} \sum_{j=1}^p \log \delta_{jd}^{(l)} \\
    & + \log K_{\frac{p+1}{2}}\(\sqrt{\eta_{d}^{(l)}
    \lambda_{\text{drug}}^{(l)}}/
    \chi_{d}^{(l)}\) + \sum_{j=1}^p \log K_{1}\(\sqrt{\delta_{jd}^{(l)}
    \lambda_{\text{feat}}^{(l)}}/\phi_{jd}^{(l)}\).
  \end{align*}

  \section{Efficient computation}\label{sec:computation}
  The empirical Bayes updates require the quantities
  $e_{jd}$, $b_{jd}$, $g_d$, and $f_d$, which in turn depend on the quantities 
  $\diag(\bSigma_d)$, $\bmu_{d}$, $\y_d \tr \X \bmu_d$, 
  $\trace(\X \tr \X \bSigma_d)$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$.
  Let $\mathbf{H}_d=g_d \diag[(b_{jd}^{(h)})^{-1}]$. Then, the first two
  quantities are efficiently calculated as:
  \begin{subequations}\label{eq:complexity1}
    \begin{align}
      \diag (\bm{\Sigma}_d^{(h+1)}) & = (a_d^{(h)})^{-1} 
      \H_d^{(h)} - (a_d^{(h)})^{-1} \{ [ \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} ] \circ
      (\H_d^{(h)} \X \tr ) \} \mathbf{1}_{n \times 1}, \\
      \bmu_d^{(h+1)}) & = \H_d^{(h)} \X \tr \y_d - \H_d^{(h)} \X \tr 
      ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr \y_d.
    \end{align}
  \end{subequations}
  The quantities $\y_d \tr \X \bmu_d$, $\bmu_d \tr \X \tr \X \bmu_d$,
  $g_d\trace[\diag(b_{jd})\bSigma_d]$, and $g_d\bmu_d \tr \diag(b_{jd}) \bmu_d$
  are easily calculated from the first two. For the remaining one we have:
  \begin{align}\label{eq:complexity2}
    \trace(\X \tr \X \bSigma_d^{(h+1)})  = 
    & (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \H_d^{(h)} \X \tr 
    \mathbf{1}_{n \times 1} \nonumber \\  
    & - (a_d^{(h)})^{-1} \mathbf{1}_{1 \times p} \{ [  \H_d^{(h)} \X \tr 
    ( \I_n + \X \H_d^{(h)} \X \tr )^{-1} \X \H_d^{(h)} \X \tr ] \circ 
    \X \tr \} \mathbf{1}_{n \times 1}.
  \end{align}
  Both (\ref{eq:complexity1}) and (\ref{eq:complexity2}) are operations of
  complexity $\mathcal{O}(pn^2)$ instead of $\mathcal{O}(p^3)$ for naive
  calculation.

@@

  We let $n_1=p/2$ and $n_2=(p-1)/2$ and use the well-known recursive relation:
  $$
  K_{\alpha}(x) = K_{\alpha - 2}(x) + \frac{2(\alpha - 1)}{2} K_{\alpha- 1}(x),
  $$
  to rewrite the ratio:
  \begin{align*}
    \frac{K_{\frac{p - 1}{2}}(x)}{K_{\frac{p + 1}{2}}(x)} = 
    \begin{cases}
      \Big( \dots \Big( \( 1 + \frac{2 \cdot 1 - 1}{x} \)^{-1} + 
      \frac{2 \cdot 2-1}{x}\Big)^{-1} + \dots + 
      \frac{2 \cdot n_1 - 1}{x}\Big)^{-1}, & \text{for } p \text{ even}, \\
      \Big( \dots \Big( \( \frac{K_0(x)}{K_1(x)} + \frac{2}{x} \cdot 1 \)^{-1} +
      \frac{2}{x} \cdot 2 \Big)^{-1} + \dots + \frac{2}{x} \cdot n_2 \Big)^{-1},
      & \text{for } p \text{ odd}.
    \end{cases}
  \end{align*} 
  The ratio $K_0(x)/K_1(x)$ is well-behaved, so that an 
  arbitrary ratio $K_{\alpha - 1}(x)/K_{\alpha}(x)$ may be calculated as
  a sequence of numerically stable scalar sums, products and inverses.

@@

  We note that through Jensen's inequality 
  $e_{jd}^{-1}=\E(\gamma_{jd}^{2})^{-1} \leq \E(\gamma_{jd}^{-2})=b_{jd}$,
  we may lowerbound the summands in left-hand side of (\ref{eq:inequality}):
  $$
  b_{jd} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} \geq
  e_{jd}^{-1} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1}.
  $$
  which we use to show that:
  \begin{align*}
    e_{jd}^{-1} + \phi_{jd}^{-2} e_{jd} - 2 \phi_{jd}^{-1} & \geq 0 \\
    1 + \phi_{jd}^{-2} e_{jd}^2 - 2 \phi_{jd}^{-1} e_{jd} & \geq 0 \\
    \( 1 - \phi_{jd}^{-1} e_{jd} \)^2 & \geq 0.
  \end{align*}
  Similar reasoning holds for the drug scale parameter $\lambda_{\text{drug}}$.

  \section{Gibbs sampler}\label{sec:gibbssampler}
  MCMC samples from the posterior corresponding to model 
  (\ref{md-eq:prior}) may be generated for each equation independently.
  MCMC samples may be generated by iteratively sampling the following 
  conditional distributions:
  \begin{subequations}\label{eq:gibbssampler}
    \begin{align}
      \bm{\beta}_d | \bgamma_d^2, \sigma_d^2, \tau_d^2, \y_d & \sim 
      \mathcal{N}_p (\bm{\mu}_d, \bm{\Sigma}_d), \label{eq:betasample} \\
      \text{with } & \bm{\Sigma}_d = \sigma_d^2 [\X \tr \X + 
      \tau_d^{-2} \diag (\gamma_{jd}^{-2})]^{-1}, \label{eq:sigmasample} \\
      \text{and } & \bm{\mu}_d = [\X \tr \X + 
      \tau_d^{-2} \diag (\gamma_{jd}^{-2})]^{-1} \X \tr 
      \y_d \label{eq:musample}, \\
      \gamma_{jd}^2 | \beta_{jd}, \sigma_d^2, \tau_d^2, \y &
      \sim \mathcal{GIG}
      \(-1, \lambda_{\text{feat}}/ \phi_{jd}^{2}, \delta_{jd}\), \\
      \text{with } & \delta_{jd} = \sigma_d^{-2} \tau_d^{-2} \beta_{jd}^2 + 
      \lambda_{\text{feat}}, \nonumber \\
      \tau_d^2 | \bm{\beta}_d, \sigma_d^2, \bgamma_{d}^2, \y & \sim
      \mathcal{GIG}
      \(-\frac{p+1}{2}, \lambda_{\text{drug}}/ \chi_{d}^{2}, \eta_d\), 
      \\
      \text{with } & \eta_d = \sigma_d^{-2} \sum_{j=1}^p \gamma_{jd}^{-2}
      \beta_{jd}^2 + \lambda_{\text{drug}}, \\
      \sigma_d^2 | \bm{\beta}_d, \bgamma_d^2, \tau_d^2, \y & \sim \Gamma^{-1} 
      \(\frac{n + p + 1}{2}, \zeta_d\), \\
      \text{with } & \zeta_d = \frac{1}{2} \[\mathbf{y}_d \tr \mathbf{y}_d -
      2 \mathbf{y}_d \tr \X \bm{\beta}_d + \bm{\beta}_d \X \tr \X \bm{\beta}_d + 
      \tau^{-2}_d \bm{\beta}_d \tr \diag(\gamma_{jd}^{-2}) \bm{\beta}_d) \]. 
    \end{align}
  \end{subequations}
  In high dimensional space, the $p \times p$ matrix inversions in 
  (\ref{eq:sigmasample}) and (\ref{eq:musample}) are   
  significant computational bottlenecks. \cite{bhattacharya_fast_2016} describe 
  a method to sample from (\ref{eq:betasample}) without explicit calculation
  of this inverse, thereby offering a siginifcant speed-up compared to naive
  sampling.

  Sampling from (\ref{eq:gibbssampler}) either requires estimating 
  (as in \ref{md-sec:empiricalbayes}) or specifying 
  hyperparameters
  $\bm{\alpha}_{\text{feat}}$, $\bm{\alpha}_{\text{drug}}$, 
  $\lambda_{\text{feat}}$, and $\lambda_{\text{drug}}$, or to endow them with
  and extra layer of hyperpriors. In general, specifying them requires 
  rarely available, detailed subject knowledge, so we may resort to endowing 
  them with hyperpriors (as in \cite{upadhyay_bayesian_1996}):
  \begin{align*}
    \bm{\alpha}_{\text{feat},g} | \lambda_{\text{feat}} & \sim 
    \mathcal{N}(0, \nu_{\text{feat}}^2/\lambda_{\text{feat}}), \\
    \bm{\alpha}_{\text{drug},h} | \lambda_{\text{drug}} & \sim 
    \mathcal{N}(0, \nu_{\text{drug}}^2/\lambda_{\text{drug}}), \\
    \lambda_{\text{feat}} & \sim \Gamma(\kappa_{\text{feat}}, 
    \xi_{\text{feat}}), \\
    \lambda_{\text{drug}} & \sim \Gamma(\kappa_{\text{drug}}, 
    \xi_{\text{drug}}),
  \end{align*}
  with hyperparameters $\nu_{\text{feat}}^2$, $\nu_{\text{drug}}^2$, 
  $\kappa_{\text{feat}}$, $\xi_{\text{feat}}$, $\kappa_{\text{drug}}$, and
  $\xi_{\text{drug}}$. Improper, flat priors occur if we let 
  $\nu_{\text{feat}}, \nu_{\text{drug}} \to \infty$ and 
  $\kappa_{\text{feat}}, \kappa_{\text{drug}}, \xi_{\text{feat}}, 
  \xi_{\text{drug}} \to 0$. The corresponding full conditionals are
  given by:
  \begin{align*}
    \bm{\alpha}_{\text{feat}} | \lambda_{\text{feat}}, \bm{\Gamma} & \sim 
    \mathcal{N} \( \mathbf{m}_{\text{feat}}, \mathbf{S}_{\text{feat}} \) \\
    \text{with } & \mathbf{S}_{\text{feat}} = \lambda_{\text{feat}}^{-1} 
    [\mathbf{C} \tr \diag(\gamma_{jd}^2) \mathbf{C} +
    \nu_{\text{feat}}^{-2} \I_G]^{-1}, \\
    \text{and } & \mathbf{m}_{\text{feat}} = [\mathbf{C} \tr 
    \diag(\gamma_{jd}^2) \mathbf{C} + \nu_{\text{feat}}^{-2} \I_G]^{-1}
    \mathbf{C} \tr \mathbf{1}_{pD \times 1}, \\
    \bm{\alpha}_{\text{drug}} | \lambda_{\text{drug}}, \bm{\tau} & \sim 
    \mathcal{N} \(\textbf{m}_{\text{drug}}, \mathbf{S}_{\text{drug}} \), \\
    \text{with } & \mathbf{S}_{\text{drug}} = \lambda_{\text{drug}}^{-1} 
    [\mathbf{Z} \tr \diag(\tau_{d}^2) \mathbf{Z} +
    \nu_{\text{drug}}^{-2} \I_H]^{-1}, \\
    \text{and } & \mathbf{m}_{\text{drug}} = [\mathbf{Z} \tr 
    \diag(\tau_{d}^2) \mathbf{Z} + \nu_{\text{drug}}^{-2} \I_H]^{-1}
    \mathbf{Z} \tr \mathbf{1}_{D \times 1}, \\
    \lambda_{\text{feat}} | \bm{\alpha}_{\text{feat}}, \bm{\Gamma} & 
    \sim \Gamma \(\frac{pD}{2} + \kappa_{\text{feat}}, 
    \upsilon_{\text{feat}}\), \\
    \text{with } & \upsilon_{\text{feat}} =\frac{1}{2} 
    \[\mathbf{C} \bm{\alpha}_{\text{feat}} - \diag(\gamma_{jd}^{-2}) 
    \mathbf{1}_{pD \times 1}\] \tr \diag(\gamma_{jd}^2) 
    \[\mathbf{C} \bm{\alpha}_{\text{feat}} - 
    \diag(\gamma_{jd}^{-2}) \mathbf{1}_{pD \times 1}\] + \xi_{\text{feat}}, \\
    \lambda_{\text{drug}} | \bm{\alpha}_{\text{drug}}, \bm{\tau} & 
    \sim \Gamma \(\frac{D}{2} + \kappa_{\text{drug}}, 
    \upsilon_{\text{drug}} \), \\
    \text{with } & \upsilon_{\text{drug}} =\frac{1}{2} 
    \[\mathbf{Z} \bm{\alpha}_{\text{drug}} - \diag(\tau_{d}^{-2}) 
    \mathbf{1}_{D \times 1}\] \tr \diag(\tau_{d}^2) 
    \[\mathbf{Z} \bm{\alpha}_{\text{drug}} - 
    \diag(\tau_{d}^{-2}) \mathbf{1}_{D \times 1}\] + \xi_{\text{drug}},
  \end{align*}
  which may be sampled together with (\ref{eq:gibbssampler}).
